### 2022-12-21 resume.qmd

---
title: "Resume"
---

[Detailed CV](Docs/JDT_CV.pdf)

### Research Interests
Statistical Image Understanding, Signal Analysis, Target Classification, Detection Theory, Computational Statistics

### Education
**Doctor of Philosophy** in Statistics (2014)		
*Florida State University, Tallahassee, FL*		
**Thesis title:** Functional Component Analysis and Regression Using Elastic Methods [PDF](Papers/dissertation.pdf)		
**Thesis advisor:** Prof. Anuj Srivastava		
**Thesis co-advisor:** Prof. Wei Wu

**Master of Science** in Electrical Engineering (2009)		
*Colorado State University, Fort Collins, CO*		
**Thesis title:** Coherence-based Underwater Target Detection for Side-Scan Sonar Imagery [PDF](Papers/MS%20Thesis.pdf)		
**Thesis advisor:** Prof. Mahmood Azimi

**Bachelor of Science** in Electrical Engineering (2007)		
*Colorado State University, Fort Collins, CO*		
**Senior project:** Magnetic Spin Wave Detection and Localization

### Experience	
**Adjunct Faculty** (2020-Present)
University of Illinois
Department of Statistics
Champaign, IL

**Distinguished Member of Technical Staff** (2022-Present)				
Sandia National Laboratories		
Statistical Sciences		
Albuquerque, NM

**Principal Member of Technical Staff** (2017-2022)				
Sandia National Laboratories		
Statistical Sciences		
Albuquerque, NM

**Senior Member of Technical Staff** (2014-2017)				
Sandia National Laboratories		
Statistical Sciences		
Albuquerque, NM

**Research Assistant** (2011-2014)		
Statistics		
Florida State University, Tallahassee, FL

**Research Engineer** (2009 - 2014)		
Naval Surface Warfare Center		
Panama City Division - Code X13		
Panama City, FL

**Research Engineer** (2007 - 2009)		
Information System Technologies, INC 		
Fort Collins, CO

**Research Assistant** (2006-2009)		
Digital Signal and Image Laboratory		
Colorado State University, Fort Collins, CO

### Affiliations	
IEEE (Senior Member), ASA, SIAM