### 2022-12-21 index.qmd
update description

---
title: "J. Derek Tucker"
description: | 
             | *Distinguished Member of Technical Staff*, Sandia National Laboratories 
             | *Adjunct Professor*, Department of Statistics, University of Illinois
image: images/me.jpg
about:
  template: trestles
  links:
    - icon: linkedin
      url: "https://linkedin.com/in/derek-tucker-04247818"
    - icon: github
      url: "https://github.com/jdtuck"
    - icon: google
      url: "https://scholar.google.com/citations?user=1jdvTZwAAAAJ&hl=en"
    - icon: file-pdf
      url: "http://arxiv.org/a/tucker_j_1"
    - icon: bi-envelope
      url: "mailto:jdtuck@sandia.gov"
    - text: CV
      href: "Docs/JDT_CV.pdf"
resources:
    - "Docs/JDT_CV.pdf"
---

::: {.column-page}

## Bio
I obtained my Ph.D. in Statistics from [Florida State University](http://www.fsu.edu/). I received the B.S. and M.S. degrees from [Colorado State University](https://www.colostate.edu/), Fort Collins, in 2007 and 2009, respectively, both in electrical engineering, with an emphasis on statistical signal processing. My dissertation was advised by [Anuj Srivastava](http://ani.stat.fsu.edu/~anuj/) and [Wei Wu](http://ani.stat.fsu.edu/~wwu/) and focused modeling of unaligned functional data. Currently, I am a Distinguished Member of the Technical Staff at Sandia National Laboratories in Albuquerque, NM and an adjunct faculty member of the Department of Statistics at the University of Illinois. At Illinois I teach a topics course on functional data analysis. Recently, I received the Director of National Intelligence Team Award for contributions to the Signal Location in Complex Environments (SLiCE) team.

My main interest lies in the area of functional data analysis with a focus on geometric methods and unaligned data. I recently have emphasized on statistical analysis and modeling of this type of data. My other research interests include statistical image understanding, statistical learning methods, and point processes.

## Education

My full CV is available [here](Docs/JDT_CV.pdf).

**Florida State University**
Tallahasee, FL

Ph.D. in Statistics | August 2011 - June 2014

**Colorado State University**
Fort Collins, CO

M.S. in Electrical Engineering | August 2007 - May 2009\
B.S. in Electrical Engineering | August 2004 - May 2007

## Contact Information

J. Derek Tucker			
[Statistical Sciences](http://www.sandia.gov/statistics/)  
[Sandia National Laboratories](http://www.sandia.gov/)  
PO Box 5800 MS 0829		
Albuquerque, NM 87185

Email SNL: [jdtuck@sandia.gov](mailto:jdtuck@sandia.gov)  
Email UIUC: [dtucker1@illinois.edu](mailto:dtucker1@illinois.edu)  
Phone: 505-284-8415

:::